\chapter*{References}
\addcontentsline{toc}{chapter}{References}

\section*{Government and Regulatory Publications}

\begin{enumerate}[label={[\arabic*]}]

    \item Reserve Bank of India. (2024). \textit{Payment System Indicators: Annual Report 2023-24}. Mumbai: RBI Press. Retrieved from https://www.rbi.org.in/

    \item Reserve Bank of India. (2022). \textit{Circular on Linking of RuPay Credit Cards to Unified Payments Interface (UPI)}. RBI/2022-23/80. Mumbai: RBI.

    \item Reserve Bank of India. (2024). \textit{Monthly Bulletin: Payment System Statistics} (April 2020 -- March 2025). Mumbai: RBI.

    \item National Payments Corporation of India. (2023). \textit{The UPI Handbook: Product Statistics and Roadmap}. Mumbai: NPCI Public Documents.

    \item National Payments Corporation of India. (2024). \textit{UPI Product Statistics} (Monthly Data 2020--2025). Retrieved from https://www.npci.org.in/

    \item NITI Aayog. (2023). \textit{Digital Payments: Trends, Issues and Opportunities}. New Delhi: Government of India.

    \item Ministry of Electronics and Information Technology. (2023). \textit{India's Digital Payment Revolution}. New Delhi: Government of India.

\end{enumerate}

\section*{Academic Journals and Books}

\begin{enumerate}[label={[\arabic*]}, resume]

    \item Davis, F. D. (1989). Perceived Usefulness, Perceived Ease of Use, and User Acceptance of Information Technology. \textit{MIS Quarterly}, 13(3), 319--340. https://doi.org/10.2307/249008

    \item Venkatesh, V., Morris, M. G., Davis, G. B., \& Davis, F. D. (2003). User Acceptance of Information Technology: Toward a Unified View. \textit{MIS Quarterly}, 27(3), 425--478. https://doi.org/10.2307/30036540

    \item Rogers, E. M. (2003). \textit{Diffusion of Innovations} (5th ed.). New York: Free Press.

    \item Agarwal, S., \& Gupta, R. (2022). The Death of Cash: Analyzing the Demonetization Effect on Digital Payments. \textit{Journal of Indian Business Research}, 14(3), 245--260.

    \item Bhattacharya, H. (2023). Credit on UPI: A Game Changer for Financial Inclusion? \textit{Economic and Political Weekly}, 58(12), 42--48.

    \item Chakravorti, B. (2022). The Digital Pivot: How India is Reshaping the Global Payment Landscape. \textit{Harvard Business Review}, 100(5), 112--119.

    \item Modi, R., \& Kumar, P. (2023). Merchant Discount Rate (MDR) and Small Retailers: A Study of Resistance. \textit{Indian Journal of Commerce and Management}, 10(2), 45--58.

    \item Duarte, F., Frost, J., Gambacorta, L., Shin, H. S., \& Wilkens, P. (2022). Instant Payments and the Demise of Cash: Evidence from Brazil's Pix. \textit{Federal Reserve Bank of Chicago Working Paper}, No. 2022-15.

    \item Chen, L., \& Zhang, Y. (2021). The Rise of QR Code Payments in China: Disruption of the Card Networks. \textit{Asian Economic Policy Review}, 16(1), 88--105.

    \item Mallat, N. (2007). Exploring Consumer Adoption of Mobile Payments: A Qualitative Study. \textit{The Journal of Strategic Information Systems}, 16(4), 413--432.

    \item Dahlberg, T., Guo, J., \& Ondrus, J. (2015). A Critical Review of Mobile Payment Research. \textit{Electronic Commerce Research and Applications}, 14(5), 265--284.

    \item Slade, E. L., Williams, M. D., \& Dwivedi, Y. K. (2014). Devising a Research Model to Examine Adoption of Mobile Payments: An Extension of UTAUT2. \textit{The Marketing Review}, 14(3), 310--335.

\end{enumerate}

\section*{Industry Reports and White Papers}

\begin{enumerate}[label={[\arabic*]}, resume]

    \item PwC India. (2024). \textit{The Indian Payments Handbook: 2024--2029}. Mumbai: PricewaterhouseCoopers.

    \item Worldline India. (2024). \textit{India Digital Payments Report: Annual Trends}. Worldline Insight Series.

    \item Boston Consulting Group \& PhonePe. (2023). \textit{Digital Payments in India: The Next Frontier}. BCG-PhonePe Pulse Report.

    \item Infosys Finacle. (2023). \textit{Banking on the Cloud: The Shift to Virtual Issuance}. Industry Whitepaper.

    \item Visa. (2023). \textit{Visa Annual Report 2023: Global Payment Trends}. San Francisco: Visa Inc.

    \item Mastercard. (2024). \textit{The Future of Payments: Digital First}. Mastercard Economics Institute.

    \item KPMG India. (2023). \textit{Fintech in India: Powering a Digital Economy}. KPMG Report.

    \item Deloitte India. (2024). \textit{The Great Payment Shift: From Plastic to Pixels}. Deloitte Insights.

\end{enumerate}

\section*{Online and Digital Sources}

\begin{enumerate}[label={[\arabic*]}, resume]

    \item NPCI Official Website. (2024). \textit{UPI Live Statistics Dashboard}. Retrieved from https://www.npci.org.in/what-we-do/upi/upi-ecosystem-statistics

    \item RBI Data Portal. (2024). \textit{Database on Indian Economy: Payment System}. Retrieved from https://dbie.rbi.org.in/

    \item Press Information Bureau. (2024). \textit{Government Initiatives on Digital Payments}. Government of India. Retrieved from https://pib.gov.in/

    \item The Economic Times. (2024). Credit on UPI: How it Works and Why it Matters. \textit{ET Tech}. Retrieved from https://economictimes.indiatimes.com/

    \item LiveMint. (2024). UPI Crosses 200 Billion Transactions: A New Milestone. \textit{Mint}. Retrieved from https://www.livemint.com/

    \item Business Standard. (2023). The Rise and Rise of UPI in India. \textit{Business Standard}. Retrieved from https://www.business-standard.com/

\end{enumerate}

\section*{Dissertations and Theses}

\begin{enumerate}[label={[\arabic*]}, resume]

    \item Sharma, A. (2022). \textit{Consumer Adoption of UPI in Urban India: An Empirical Study} (Unpublished Master's Dissertation). Delhi School of Economics, University of Delhi.

    \item Verma, P. (2023). \textit{Digital Payment Adoption Among Small Merchants: A Study of Delhi-NCR} (Unpublished MBA Project). Faculty of Management Studies, University of Delhi.

\end{enumerate}
